---
title: "01_geocode_data"
format: html
editor: source
---

# Setup

```{r}
#| echo: false
#| warning: false
library(tidyverse)
library(sf)
library(data.table)
library(janitor)
library(fs)
library(ggmap)
```

# Read data

```{r}
df_montreal <- data.table::fread("data/MontrealRawData_2023_08_24.csv")
df_montreal <- janitor::clean_names(df_montreal)
```


# Geolocate 

Some observations don't have the necessary location longitude and latitude. We need to geocode them.

However, some of these observations also don't have the necessary information for geocoding (either the postal code, at `q51`, or the nearest intersection, at `q50`). 
We will, then, geocode the observations that we can, i.e., those we have available information.

```{r}
#| label: fix postal codes
data.table::setDT(df_montreal)

df_montreal[nchar(q51)==10, q51 := gsub("[[:space:]]", "", q51)]

df_montreal[
  ,
  q51 := str_trim(q51)
]

df_montreal[
  nchar(q51)==6
  , q51 := gsub(q51, pattern = "(.{3})(.*)", replacement = "\\1 \\2")
]


df_montreal[q51=="H4x c1c7", q51 := "H4x 1c7"]
df_montreal[q51=="H4G 2P2’", q51 := "H4G 2P2"]

```

```{r}
#| label: geocode
key <- data.table::fread(file = "data/google_key.csv", header = F)

ggmap::register_google(key = key)

# geocode intersections
intersections <- df_montreal %>% filter(is.na(location_latitude) & nchar(q50)!= 0) %>% 
  select(v1, q50)

geo1 <- purrr::map(
  intersections$q50
  , ~ggmap::geocode(
    ., output = "latlona", source = "google"
  )
)

# error with intersections #2 ("Laurier et Parc")
# remove from object
geo1 <- geo1[[1]]

# add v1 for referencing
geo1$v1 <- intersections$v1[[1]]

# geocode postal codes
postal <- df_montreal %>% 
  filter(is.na(location_latitude) & nchar(q51)!=0) %>% 
  select(v1, q51)

geo2 <- purrr::map(
  postal$q51
  , ~ggmap::geocode(
    .
    , output = "latlona", source = "google"
  )
)

geo2 <- data.table::rbindlist(geo2)

# add v1 for referencing
geo2$v1 <- postal$v1

geo <- rbindlist(list(geo1, geo2))
data.table::setDT(geo)

geo <- geo[order(v1)]
df_montreal <- df_montreal[order(v1)]

df_montreal[
  v1 %in% geo$v1
  , `:=`(
      location_latitude = geo$lat
      , location_longitude = geo$lon
  )
]
```

# Filter 
Filter observations with available location data.

```{r}
df_montreal_filter <- df_montreal %>% 
  filter(!is.na(location_latitude))
```


# Convert to sf
```{r}
df_montreal_filter_sf <- df_montreal_filter %>% 
  sf::st_as_sf(coords = c("location_longitude","location_latitude"))

# set crs 4326
sf::st_crs(df_montreal_filter_sf) <- 4326
```


# Save data

```{r}
data.table::fwrite(df_montreal, file = "data/df_survey_montreal_geolocated_complete.csv")
data.table::fwrite(df_montreal_filter, file = "data/df_survey_montreal_geolocated_filtered.csv")
sf::st_write(df_montreal_filter_sf, "data/df_survey_montreal_filter_sf.gpkg", append = F)
```



