---
title: "Mobilizing Justice: Montreal Fair Pass Pilot"
author:
  - name: João Pedro Figueira Amorim Parga
    orcid: 0000-0002-4105-5927
    email: joao.parga@mail.utoronto.ca
    affiliations:
      - name: Mobilizing Justice Partnership
format:
  html:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: true
    code_folding: hide
    theme: readable
    embed-resources: true
  pdf:
    number-sections: true
    colorlinks: true
    keeptex: true
    include-in-header: 
      text: |
        \usepackage{booktabs}
        \usepackage{siunitx}
        \newcolumntype{d}{S[
            input-open-uncertainty=,
            input-close-uncertainty=,
            parse-numbers = false,
            table-align-text-pre=false,
            table-align-text-post=false
         ]}
date: 'last-modified'
date-format: '[This version:] MMMM D, YYYY'
abstract: 'This report provides an overview of the first wave of the Montreal Fair Pass Survey.'
bibliography: bib/zoterolibrary_updated_10_05_23.bib
csl: csl/american-sociological-association.csl
tbl-cap-location: top
number-sections: true
execute:
  echo: false
  warning: false
  message: false
  cache: false
editor:
  mode: source
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)
# Use cache = TRUE if you want to speed up compilation

knitr::opts_knit$set(output.format = "html")  # Set to "html" for HTML output

# A function to allow for showing some of the inline code
rinline <- function(code){
  html <- '<code  class="r">``` `r CODE` ```</code>'
  sub("CODE", code, html)
}
```

```{r packages}
#| echo: false
#| include: false
library(tidyverse)
library(sf)
library(data.table)
library(janitor)
library(fs)
library(cancensus)
library(tmap)
library(DataExplorer)
library(patchwork)
library(mapview)
```


<!-- READ AND MANIPULATE DATA -->

```{r readdata}
#| include: false
# montreal fsa
montreal_fsa <- sf::st_read("data/report/montreal_fsa_count.gpkg")
# montreal cd (ct)
montreal_cd_ct <- sf::st_read("data/report/montreal_cd_ct_count.gpkg")
# survey raw data
survey_raw <- data.table::fread("data/utscfiletransfer/Montreal_Corrected_Geocoding_2023_09_18.csv")
survey_raw <- janitor::clean_names(survey_raw)
# survey filtered data
survey_filter_cd <- sf::st_read("data/report/survey_filter_inside_cd.gpkg")
```

```{r createvars}
survey_filter_cd <- survey_filter_cd %>% 
  mutate(
    income_midpoint = case_when(
      q31 == "Less than $15,000" ~ 15000/2
      , q31 == "$15,000 to $29,999 " ~ (15000 + 30000)/2
      , q31 == "$30,000 to $59,999" ~ (30000 + 60000)/2
      , q31 == "$60,000 to $89,999 " ~ (60000 + 90000)/2
      , q31 == "$90,000 to $119,999 " ~ (90000 + 120000)/2
      , q31 == "$120,000 to $149,999 " ~ (120000 + 150000)/2
      , q31 == "$150,000 to $179,999 " ~ (150000 + 180000)/2
      , q31 == "$180,000 to $209,999 " ~ (180000 + 210000)/2
      , q31 == "$210,000 or more" ~ 210000
      , .default = NA_integer_
    )
    , household_size = case_when(
      q106 == "1" ~ 1
      , q106 == "2" ~ 2
      , q106 == "3" ~ 3
      , q106 == "4" ~ 4
      , q106 == "5" ~ 5
      , q106 == "6" ~ 6
      , q106 == "8" ~ 8
      , q106 == "10 or more" ~ 10
      , .default = NA_integer_
    )
    , gender = case_when(
      q41 == "Man" ~ "male"
      , q41 == "Woman" ~ "female"
      , q41 %in% c("Non-binary", "Another gender identity, please specify:") ~ 
        "other"
      , .default = NA_character_
    )
    , age = case_when(
      q40 == "1923 or earlier" ~ 2023L - 1923L
      ,q40 == "1930" ~ 2023L - 1930L
      ,q40 == "1935" ~ 2023L - 1935L
      ,q40 == "1937" ~ 2023L - 1937L
      ,q40 == "1938" ~ 2023L - 1938L
      ,q40 == "1939" ~ 2023L - 1939L
      ,q40 == "1940" ~ 2023L - 1940L
      ,q40 == "1941" ~ 2023L - 1941L
      ,q40 == "1942" ~ 2023L - 1942L
      ,q40 == "1943" ~ 2023L - 1943L
      ,q40 == "1944" ~ 2023L - 1944L
      ,q40 == "1945" ~ 2023L - 1945L
      ,q40 == "1946" ~ 2023L - 1946L
      ,q40 == "1947" ~ 2023L - 1947L
      ,q40 == "1948" ~ 2023L - 1948L
      ,q40 == "1949" ~ 2023L - 1949L
      ,q40 == "1950" ~ 2023L - 1950L
      ,q40 == "1951" ~ 2023L - 1951L
      ,q40 == "1952" ~ 2023L - 1952L
      ,q40 == "1953" ~ 2023L - 1953L
      ,q40 == "1954" ~ 2023L - 1954L
      ,q40 == "1955" ~ 2023L - 1923L
      ,q40 == "1956" ~ 2023L - 1956L
      ,q40 == "1957" ~ 2023L - 1957L
      ,q40 == "1958" ~ 2023L - 1958L
      ,q40 == "1959" ~ 2023L - 1959L
      ,q40 == "1960" ~ 2023L - 1960L
      ,q40 == "1961" ~ 2023L - 1961L
      ,q40 == "1962" ~ 2023L - 1962L
      ,q40 == "1963" ~ 2023L - 1963L
      ,q40 == "1963" ~ 2023L - 1963L
      ,q40 == "1963" ~ 2023L - 1963L
      ,q40 == "1963" ~ 2023L - 1963L
      ,q40 == "1964" ~ 2023L - 1964L
      ,q40 == "1965" ~ 2023L - 1965L
      ,q40 == "1966" ~ 2023L - 1966L
      ,q40 == "1967" ~ 2023L - 1967L
      ,q40 == "1968" ~ 2023L - 1968L
      ,q40 == "1969" ~ 2023L - 1969L
      ,q40 == "1970" ~ 2023L - 1970L
      ,q40 == "1971" ~ 2023L - 1971L
      ,q40 == "1972" ~ 2023L - 1972L
      ,q40 == "1973" ~ 2023L - 1973L
      ,q40 == "1974 or later" ~ 2023L - 1974L
      , .default = NA_integer_
    )
    # visible minority 
    # https://www23.statcan.gc.ca/imdb/p3Var.pl?Function=DECI&Id=257515
    , visible_minority = case_when(
      stringr::str_detect(
        string = q48
        , pattern = "Southeast Asian origins|South Asian origins|Northeast Asian origins|West Asian origins|Caribbean origins|Latin, Central and South American origins|African origins|Arabic origins/North African origins") ~ "visible_minority"
      , stringr::str_detect(
        string = q48
        , pattern = "Native North American origins"
      ) ~ "native_origins"
      , stringr::str_detect(
        string = q48
        , pattern = "I prefer not to answer"
      ) ~ NA_character_
      , .default = "not_visible_mninority"
    )
  )

survey_filter_cd <- survey_filter_cd %>% 
  mutate(
    # https://www12.statcan.gc.ca/census-recensement/2021/ref/dict/tab/index-eng.cfm?ID=t2_4
    lim_threshold = case_when(
      household_size == 1L ~ 26503
      , household_size == 2L ~ 37480
      , household_size == 3L ~ 45904
      , household_size == 4L ~ 53005
      , household_size == 5L ~ 59261
      , household_size == 6L ~ 64918
      , household_size == 8L ~ 74962
      , household_size == 10L ~ 83810
      , .default = NA_integer_
    )
  )

survey_filter_cd <- survey_filter_cd %>% 
  mutate(
    # https://www12.statcan.gc.ca/census-recensement/2021/ref/dict/tab/index-eng.cfm?ID=t2_4
    in_lim = case_when(
      income_midpoint < lim_threshold ~ "yes"
      , income_midpoint >= lim_threshold ~ "no"
      , .default = NA_character_
    )
  )
```



<!-- START OF THE DOCUMENT -->

# Initial information

The raw dataset contained `r 2 +2` observations.

# Spatial distribution {#sec-spatial}

Below, we plot the spatial distribution of the sample against the spatial distribution of the population over 50 years old, according to the 2021 Census.


```{r, fig.align="center", fig.pos="H"}
#| label: map-1-under-over-fsa
#| fig-cap: "Under/over-representation: Ratio of the proportions of survey respondents vs. the population over 50 in each FSA (left)"
#| out-width: 100%

tm_shape(montreal_fsa) + 
  tm_polygons(
    col = "under_over"
    , breaks = c(0,0.5, 0.99, 1.01, 2, Inf)
    , palette = "BrBG"
    , border.alpha = 0.25
    , alpha = 0.95
    , labels = c("0.00-0.50: Survey underrepresents","0.50-0.99", '0.99-1.01', '1.01-2.00', '2.00 or more: Survey overrepresents')
    , title = "Ratio: Survey/(Census)"
    ) + 
  tm_text(
    text = "label_survey"
    , size = 0.3
    # , col = "white"
    ) + 
  tm_shape(montreal_fsa) +
  tm_text(
      text = "label_census"
      , size = 0.3
      , ymod = -0.25
      # , col = "black"
  ) +
  tm_add_legend(
    type = "text"
  ) +
  tm_layout(
    legend.width = 2
    # , main.title = "Ratio of the proportions of survey respondents vs. the population over 50 in each CT"
    # , main.title.position = "center"
    # , main.title.size = 1.05
    , legend.title.size = 1.2
    , legend.text.size = 0.8
  )

# tmap_arrange(tm_under_over_fsa, tm_under_over_ct)
```
```{r, fig.align="center", fig.pos="H"}
#| label: map-2-under-over-ct
#| fig-cap: "Under/over-representation: Ratio of the proportions of survey respondents vs. the population over 50 in each CT"
#| out-width: 100%

# tm_under_over_ct <- 
  tm_shape(montreal_cd_ct) + 
  tm_polygons(
    col = "under_over"
    , breaks = c(0,0.5, 0.99, 1.01, 2, Inf)
    , palette = "BrBG"
    , border.alpha = 0.5
    , alpha = 0.95
    , labels = c("0.00-0.50: Survey underrepresents","0.50-0.99", '0.99-1.01', '1.01-2.00', '2.00 or more: Survey overrepresents')
    , title = "Ratio: Survey/(Census)"
  ) + 
  tm_add_legend(
    type = "text"
  ) +
  tm_layout(
    legend.width = 2
    # , main.title = "Ratio of the proportions of survey respondents vs. the population over 50 in each CT"
    # , main.title.position = "center"
    # , main.title.size = 1.05
    , legend.title.size = 1.2
    , legend.text.size = 0.8
  )

# tmap_arrange(tm_under_over_fsa, tm_under_over_ct)
```


\newpage

# References {.unnumbered}

::: {#refs}
:::

<!--  to generate a citation entry for BibTeX, you can pass the returned object of citation() to toBibtex(), e.g. toBibtex(citation("xaringan")). Then  copy the output to a .bib file and add a unique citation key -->

<!-- https://bookdown.org/yihui/rmarkdown-cookbook/write-bib.html -->

\newpage

# Online appendix {.unnumbered}

## Attach R session info in appendix {.unnumbered}

R session version and package information.

```{r echo=FALSE}
print(sessionInfo(), local = FALSE)
```

## All the code in the paper {.unnumbered}

Below, all the code used in the file:

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```

