---
title: "Mobilizing Justice: Montreal Fair Pass Pilot"
author:
  - name: João Pedro Figueira Amorim Parga
    orcid: 0000-0002-4105-5927
    email: joao.parga@mail.utoronto.ca
    affiliations:
      - name: Mobilizing Justice Partnership
format:
  html:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: true
    code_folding: hide
    theme: readable
    embed-resources: true
  pdf:
    fig-width: 6
    fig-asp: 0.618
    number-sections: true
    colorlinks: true
    keeptex: true
    include-in-header: 
      text: |
        \usepackage{booktabs}
        \usepackage{siunitx}
        \newcolumntype{d}{S[
            input-open-uncertainty=,
            input-close-uncertainty=,
            parse-numbers = false,
            table-align-text-pre=false,
            table-align-text-post=false
         ]}
date: 'last-modified'
date-format: '[This version:] MMMM D, YYYY'
abstract: 'This report provides an overview of the first wave of the Montreal Fair Pass Survey.'
bibliography: bib/zoterolibrary_updated_10_05_23.bib
csl: csl/american-sociological-association.csl
tbl-cap-location: top
number-sections: true
execute:
  echo: false
  warning: false
  message: false
  cache: false
editor:
  mode: source
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)
# Use cache = TRUE if you want to speed up compilation

knitr::opts_knit$set(output.format = "html")  # Set to "html" for HTML output

# A function to allow for showing some of the inline code
rinline <- function(code){
  html <- '<code  class="r">``` `r CODE` ```</code>'
  sub("CODE", code, html)
}
```

```{r packages}
#| echo: false
#| include: false
library(tidyverse)
library(sf)
library(data.table)
library(janitor)
library(fs)
library(cancensus)
library(tmap)
library(DataExplorer)
library(patchwork)
library(mapview)
library(modelsummary)
library(kableExtra)
library(knitr)
library(gt)
library(conflicted)
library(stars)
library(showtext)
library(sysfonts)
```

```{r conflicts}
#| include: false
select <- dplyr::select
conflicted::conflicts_prefer(dplyr::filter)
```

```{r fonts}
#| include: false
#| # use different fonts
sysfonts::font_add_google("Roboto", "roboto")
showtext::showtext_auto()
```


<!-- READ AND MANIPULATE DATA -->

```{r readdata}
#| include: false
# montreal fsa
montreal_fsa <- sf::st_read("data/report/montreal_fsa_count.gpkg")
# montreal cd (ct)
montreal_cd_ct <- sf::st_read("data/report/montreal_cd_ct_count.gpkg")
# survey raw data
n_obs_survey_raw <- data.table::fread("data/utscfiletransfer/Montreal_Corrected_Geocoding_2023_09_18.csv") %>% 
  nrow()
# survey geocoded
n_obs_survey_geocoded <- 
  sf::st_read("data/df_survey_montreal_filter_sf_new.gpkg") %>% 
  nrow()
# survey filtered data
survey_filter_cd <- sf::st_read("data/report/survey_filter_inside_cd.gpkg")
```




<!-- START OF THE DOCUMENT -->

# Initial information

The original raw dataset contained `r n_obs_survey_raw` respondents, of which we were able to geocode `r n_obs_survey_geocoded` observations.
`r n_obs_survey_geocoded - nrow(survey_filter_cd)` of them declared their residential location outside a 250 meter buffer around the island of Montreal (`CD == 2466`), leaving us with `r nrow(survey_filter_cd)` observations in the final dataset.

# Spatial distribution {#sec-spatial}

```{r}
#| include: false
fsa <- montreal_fsa %>% 
  select(geouid) %>% 
  rename("fsa_geouid" = "geouid")
ct <- montreal_cd_ct %>% 
  select(geo_uid, count_lim_tot_parent, count_lim_tot_in_lim) %>% 
  rename("ct_geo_uid" = "geo_uid")

fsa_ct_join <- sf::st_join(fsa, ct)

total_lim_sum <- sum(montreal_cd_ct$count_lim_tot_parent)

fsa_ct_grouped <- fsa_ct_join %>% 
  group_by(fsa_geouid) %>% 
  summarise(
    sum_lim_tot_parent = sum(count_lim_tot_parent, na.rm = T)
    , sum_lim_tot_in_lim = sum(count_lim_tot_in_lim, na.rm = T)
    , prop_tot_pop = sum_lim_tot_parent / total_lim_sum
    , prop_tot_in_lim = sum_lim_tot_in_lim / total_lim_sum
  )
```


## Over 50 years 
Below, we plot the spatial distribution of the sample against the spatial distribution of the population over 50 years old, according to the 2021 Census.
We see that the survey over represents on the eastern part of the island, and somewhat in its center. 
The CTs that are under represented are located in the western part.

```{r, fig.align="center", fig.pos="H", fig.cap=c("Population in LIM (Census)", "Under/over-representation ratio of survey respondents vs. population over 50")}
#| label: map-1-under-over-fsa
#| out-width: 100%
#| fig-format: pdf
#| fig-asp: 1
#| layout-ncol: 2

# quarto figures
# https://r4ds.hadley.nz/quarto.html
# plot scaling
# https://quarto.org/docs/reference/cells/cells-knitr.html
# multiple plots and captions
# https://stackoverflow.com/a/41231696

# fig-cap: "Under/over-representation: Ratio of the proportions of survey respondents vs. the population over 50 in each FSA (left)"

tm_shape(fsa_ct_grouped) +
  tm_polygons(
    col = "prop_tot_in_lim"
    , style = "jenks"
    , palette = "YlGnBu"
    , border.alpha = 0.5
    , alpha = 0.95
    , title = "Proportion of Residents in LIM (relative to total population)"
  ) + 
  tm_layout(
    legend.width = 2
    # , main.title = "Ratio of the proportions of survey respondents vs. the population over 50 in each CT"
    # , main.title.position = "center"
    # , main.title.size = 1.05
    , legend.title.size = 1.5
    , legend.text.size = 1
    , fontfamily = "roboto"
  ) +
  tm_compass(color.dark = "grey30", text.color = "grey30") +
  tm_scale_bar(
    breaks = c(0,2)
    , color.dark = "grey30"
    , text.color = "grey30"
    , position = "left"
    )

tm_shape(montreal_fsa) + 
  tm_polygons(
    col = "under_over"
    , breaks = c(0,0.5, 0.99, 1.01, 2, Inf)
    , palette = "RdBu"
    , border.alpha = 0.25
    , alpha = 0.95
    , labels = c("0.00-0.50: Survey underrepresents","0.50-0.99", '0.99-1.01', '1.01-2.00', '2.00 or more: Survey overrepresents')
    , title = "Ratio: Survey/(Census)"
    ) + 
  tm_text(
    text = "label_survey"
    , size = 0.3
    # , col = "white"
    ) + 
  tm_shape(montreal_fsa) +
  tm_text(
      text = "label_census"
      , size = 0.3
      , ymod = -0.25
      # , col = "black"
  ) +
  tm_add_legend(
    type = "text"
  ) +
  tm_layout(
    legend.width = 2
    # , main.title = "Ratio of the proportions of survey respondents vs. the population over 50 in each CT"
    # , main.title.position = "center"
    # , main.title.size = 1.05
    , legend.title.size = 1.2
    , legend.text.size = 0.8
    , fontfamily = "roboto"
  ) +
  tm_compass(color.dark = "grey30", text.color = "grey30") +
  tm_scale_bar(breaks = c(0,2), color.dark = "grey30", text.color = "grey30"
               , position = "left")

# tmap_arrange(tm_under_over_fsa, tm_under_over_ct)
```



# Table

