---
title: "Mobilizing Justice: Montreal Fair Pass Pilot"
author:
  - name: João Pedro Figueira Amorim Parga
    orcid: 0000-0002-4105-5927
    email: joao.parga@mail.utoronto.ca
    affiliations:
      - name: Mobilizing Justice Partnership
format:
  docx: 
    reference-doc: custom-reference-doc.docx
    toc: false
    toc-depth: 2
    number-sections: true
  html:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: true
    code_folding: hide
    theme: readable
    embed-resources: true
  pdf:
    fig-width: 6
    fig-asp: 0.618
    number-sections: true
    colorlinks: true
    keeptex: true
    include-in-header: 
      text: |
        \usepackage{booktabs}
        \usepackage{siunitx}
        \newcolumntype{d}{S[
            input-open-uncertainty=,
            input-close-uncertainty=,
            parse-numbers = false,
            table-align-text-pre=false,
            table-align-text-post=false
         ]}
date: "last-modified"
date-format: '[This version:] MMMM D, YYYY'
abstract: "This version contains only the updated tables and visualizations from the new report. The text can be found at the onedrive link."
csl: csl/american-sociological-association.csl
tbl-cap-location: top
number-sections: true
execute:
  echo: false
  warning: false
  message: false
  cache: false
editor:
  mode: source
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)
# Use cache = TRUE if you want to speed up compilation

knitr::opts_knit$set(output.format = "html")  # Set to "html" for HTML output

# A function to allow for showing some of the inline code
rinline <- function(code){
  html <- '<code  class="r">``` `r CODE` ```</code>'
  sub("CODE", code, html)
}
```

```{r packages}
#| echo: false
#| include: false
library(tidyverse)
library(sf)
library(data.table)
library(janitor)
library(fs)
library(cancensus)
library(tmap)
library(DataExplorer)
library(patchwork)
library(mapview)
library(modelsummary)
library(kableExtra)
library(knitr)
library(gt)
library(gtsummary)
library(conflicted)
library(stars)
library(showtext)
library(sysfonts)
library(hrbrthemes)
library(flextable)
```

```{r conflicts}
#| include: false
select <- dplyr::select
conflicted::conflicts_prefer(dplyr::filter)
```

```{r fonts}
#| include: false
#| # use different fonts
sysfonts::font_add_google("Roboto", "roboto")
showtext::showtext_auto()
```


```{r}
knitr::opts_chunk$set(tab.topcaption = TRUE)
```


<!-- READ AND MANIPULATE DATA -->

```{r readdata}
#| include: false

load(file = "data/df_both_waves.RData")
load(file = "data/list_count_trips_wave.RData")

```



<!-- START OF THE DOCUMENT -->

# Descriptive analysis


## Who took the survey?

### Table response rate

```{r tbldescriptiveresponserateage, tab.id="descriptive", tab.cap="Response rate between first and second wave"}
#| echo: false
#| tbl-cap: Response rate between first and second wave

df_both %>% 
  filter(wave=="First") %>% 
  select(
 response_rate,age_groups_control_treat, age_groups_5_all
) %>% 
  rename(
    'Age groups (Control and Treatment)' = age_groups_control_treat
    , "Age groups (5 years intervals)" = age_groups_5_all
  ) %>% 
  mutate(
    `Age groups (5 years intervals)` = fct_recode(
      `Age groups (5 years intervals)`
      , `Age 50-54` = "age_50_54"
      , `Age 55-59` = "age_55_59"
      , `Age 60-64` = "age_60_64"
      , `Age 65-69` = "age_65_69"
      , `Age 70-74` = "age_70_74"
      , `Age 75-79` = "age_75_79"
      , `Age 80-84` = "age_80_84"
      , `Age 85-89` = "age_85_89"
      , `Age 90 and over` = "age_90_over"
      )
  ) %>% 
  st_drop_geometry() %>% 
  gtsummary::tbl_summary(
    type = gtsummary::all_dichotomous() ~ "categorical"
    , by = response_rate
    , sort = list(
      `Age groups (5 years intervals)` ~ "alphanumeric"
      , `Age groups (Control and Treatment)` ~ "frequency"
    ) 
    # all_categorical() ~ "alphanumeric"
    , missing_text = "Missing"
  ) %>% 
  gtsummary::modify_caption(caption = "Age groups of the respondents") %>% 
  as_flex_table() %>% 
  flextable::font(fontname = "Arial")

# gtsummary::as_gt() %>% 
# gt::tab_options(table.font.names = "Arial")
#%>% 
# as_flex_table() 
#%>% 
# gtsummary::modify_header(label = "Demographics") %>% 
# gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables")
```

```{r tbldescriptiveresponserateinvariable, tab.id="descriptive", tab.cap="Response rate between first and second wave"}
#| echo: false
#| tbl-cap: Response rate between first and second wave

df_both %>% 
  filter(wave == "First") %>% 
  select(
    response_rate
  , visible_minority
  , immigrant
  , gender
  , age_groups_control_treat
  , educ_has_postsec
) %>% 
  rename(
    "Visible minority" = visible_minority
    , "Is Immigrant" = immigrant
    , "Gender" = gender
    , 'Age Groups Control Treat' = age_groups_control_treat
    , "Has post-secondary education" = educ_has_postsec
  ) %>% 
  tbl_strata(
    strata = response_rate
    , .header = "**{strata}**"
    , ~.x %>% 
      gtsummary::tbl_summary(
        type = gtsummary::all_dichotomous() ~ "categorical"
        , sort = list(
          `Visible minority` ~ "frequency"
          , `Is Immigrant` ~ "frequency"
          , `Gender` ~ "frequency"
          , `Has post-secondary education` ~ "frequency"
        )
        # all_categorical() ~ "frequency"
        , missing_text = "Missing"
        , by = `Age Groups Control Treat`
      ) %>% 
      gtsummary::add_overall(last = T, col_label = "**Total**, N = {N}") %>%
      gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables")
  )  %>% 
  as_flex_table() %>% 
  flextable::font(fontname = "Arial")

# gtsummary::as_gt() %>% 
# gt::tab_options(table.font.names = "Arial")
#%>% 
# as_flex_table() 
#%>% 
# gtsummary::modify_header(label = "Demographics") %>% 
# gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables")
```

### Table age

This subsection contains a general description of who took the first wave of the survey...

```{r tbldescriptiveage, tab.id="descriptive", tab.cap="Descriptive statistics for categorical age groups"}
#| echo: false
#| tbl-cap: Descriptive statistics for categorical variables

df_both %>% select(
 age_groups_control_treat, age_groups_5_all
) %>% 
  rename(
    'Age groups (Control and Treatment)' = age_groups_control_treat
    , "Age groups (5 years intervals)" = age_groups_5_all
  ) %>% 
  mutate(
    `Age groups (5 years intervals)` = fct_recode(
      `Age groups (5 years intervals)`
      , `Age 50-54` = "age_50_54"
      , `Age 55-59` = "age_55_59"
      , `Age 60-64` = "age_60_64"
      , `Age 65-69` = "age_65_69"
      , `Age 70-74` = "age_70_74"
      , `Age 75-79` = "age_75_79"
      , `Age 80-84` = "age_80_84"
      , `Age 85-89` = "age_85_89"
      , `Age 90 and over` = "age_90_over"
      )
  ) %>% 
  st_drop_geometry() %>% 
  gtsummary::tbl_summary(
    type = gtsummary::all_dichotomous() ~ "categorical"
    , sort = list(
      `Age groups (5 years intervals)` ~ "alphanumeric"
      , `Age groups (Control and Treatment)` ~ "frequency"
    ) 
    # all_categorical() ~ "alphanumeric"
    , missing_text = "Missing"
  ) %>% 
  gtsummary::modify_caption(caption = "Age groups of the respondents") %>% 
  as_flex_table() %>% 
  flextable::font(fontname = "Arial")

  # gtsummary::as_gt() %>% 
  # gt::tab_options(table.font.names = "Arial")
#%>% 
  # as_flex_table() 
#%>% 
  # gtsummary::modify_header(label = "Demographics") %>% 
  # gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables")
```
### Table invariable characteristics through time

```{r tbldescriptiveinvariable, tab.id="descriptive", tab.cap="Descriptive statistics for categorical variables constant over time"}
#| echo: false
#| tbl-cap: Descriptive statistics for categorical variables constant over time

df_both %>% filter(wave=="Second") %>% 
  select(
  visible_minority
  , immigrant
  , gender
  , age_groups_control_treat
  , educ_has_postsec
) %>% 
  rename(
    "Visible minority" = visible_minority
    , "Is Immigrant" = immigrant
    , "Gender" = gender
    , 'Age Groups Control Treat' = age_groups_control_treat
    , "Has post-secondary education" = educ_has_postsec
  ) %>% 
  gtsummary::tbl_summary(
    type = gtsummary::all_dichotomous() ~ "categorical"
    , sort = list(
      `Visible minority` ~ "frequency"
      , `Is Immigrant` ~ "frequency"
      , `Gender` ~ "frequency"
      , `Has post-secondary education` ~ "frequency"
    )
      # all_categorical() ~ "frequency"
    , missing_text = "Missing"
    , by = `Age Groups Control Treat`
  ) %>% 
  gtsummary::add_overall(last = T, col_label = "**Total**, N = {N}") %>%
  gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables") %>% 
  as_flex_table() %>% 
  flextable::font(fontname = "Arial")

  # gtsummary::as_gt() %>% 
  # gt::tab_options(table.font.names = "Arial")
#%>% 
  # as_flex_table() 
#%>% 
  # gtsummary::modify_header(label = "Demographics") %>% 
  # gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables")
```




### Table VARIABLE demographics through time


```{r tbldescriptivevariable, tab.id="descriptive", tab.cap="Descriptive statistics for categorical variables variable that vary over time"}
#| echo: false
#| tbl-cap: Descriptive statistics for categorical variables variable that vary over time

df_both %>% select(
  wave
  , age_groups_control_treat
  , any_mobility_limitation
  , any_functional_limitation
  , employment
  , income_groups
  , in_lim
  , lone_household
  , has_vehicle_household
) %>% 
  rename(
    'Age Groups Control Treat' = age_groups_control_treat
    # , "Age groups" = age_groups_5_all
    , "Any mobility limitation" = any_mobility_limitation
    , "Any functional limitation" = any_functional_limitation
    , "Employment" = employment
    , "Annual income groups" = income_groups
    , "Household below LIM" = in_lim
    , "Single-person household" = lone_household
    , "Has Vehicle in Household" = has_vehicle_household
  ) %>% 
  tbl_strata(
    strata = wave
    , .header = "**{strata}**"
    , ~.x %>% 
      gtsummary::tbl_summary(
        type = gtsummary::all_dichotomous() ~ "categorical"
        , sort = list(
          `Any mobility limitation` ~ "frequency"
          , `Any functional limitation` ~ "frequency"
          , `Employment` ~ "frequency"
          , `Annual income groups` ~ "alphanumeric"
          , `Household below LIM` ~ "frequency"
          , `Single-person household` ~ "frequency"
          , `Has Vehicle in Household` ~ "frequency"
        )
        # all_categorical() ~ "frequency"
        , missing_text = "Missing"
        , by = `Age Groups Control Treat`
      ) %>% 
      gtsummary::add_overall(last = T, col_label = "**Total**, N = {N}") %>%
      gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables")
  )  %>% 
  as_flex_table() %>% 
  flextable::font(fontname = "Arial")

# gtsummary::as_gt() %>% 
# gt::tab_options(table.font.names = "Arial")
#%>% 
# as_flex_table() 
#%>% 
# gtsummary::modify_header(label = "Demographics") %>% 
# gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables")
```

, any_mobility_limitation
  , any_functional_limitation
  , employment
  , income_groups
  , in_lim
  , lone_household
  

## What is their travel behavior?


, has_vehicle_household


### Transit use, cost barriers, and suppressed demand

The table below displays the cross-tabulations with respondents’ travel behavior by their age groups, i.e., over 65 years old (Treatment) and between 50 and 64 years old (Control). Seventy percent of the individuals over 65 years old have the Free 65+ fare loaded into their Opus card. The relative percentage of people that declared riding transit as much as they liked was higher in the treatment group (73%) than in the control (66%). Nonetheless, 85% of the treatment group said that the cost of transit has prevented them from taking a trip using public transit, compared to 66% of the control. Additionally, the treatment group reported a higher percentage (73%) of respondents that declared having difficulty meeting transport expenses when compared to the control (52%). These results deserve further investigation: is suppressed demand an issue disproportionately affecting people that do not have access to the fare-free program? And how will these results change with time after the implementation of the fare-free policy? 

Regarding suppressed transit demand because of transit costs, control and treatment behavior are similar. Active modes of transportation are the most common form of replacement for those who reported not taking a transit trip because of transit cost - more than 40% for both control and treatment. Not taking the trip at all is the second most common option (27% for the whole sample), followed by using the car (24%) as the alternative trip modal choice. 

```{r dfrename}
#| echo: false
df_rename <- df_both 
data.table::setDT(df_rename)
df_rename[
  age_groups_control_treat == "Control (50-64)"
  , has_free_pass_wave1 := NA_character_
]

names(df_rename) <- gsub("_", " ", colnames(df_both)) %>% 
  str_to_title()
df_rename <- df_rename %>% st_drop_geometry()
```



# Update

## Old analysis (1st wave only (demographics?))

```{r}
# loaded free pass on card (treat and control)
df_both %>% filter(wave=="Second") %>%  count(q111, age_groups_control_treat)
```

df_both %>% filter(wave=="Second") %>%  count(q111, age_groups_control_treat)


## Update old (1st and 2nd wave combined)

Awareness of free pass and how did that translate to its use
```{r}
# 1st wave
df_both %>% filter(wave=="Second") %>%  count(awereness_free_pass, age_groups_control_treat)
df_both %>% filter(wave=="Second") %>%  count(has_free_pass_wave1, age_groups_control_treat)
df_both %>% filter(wave=="Second") %>%  count(has_free_pass_wave1_categories, age_groups_control_treat)

# type of fare (available for 1st wave only)
df_both %>% filter(wave=="Second" & has_free_pass_wave1 == "Yes") %>%  count(fare_type_before_pass_q98, age_groups_control_treat)
df_both %>% filter(wave=="Second" & has_free_pass_wave1 == "Yes") %>%  count(fare_type_before_pass_q99, age_groups_control_treat)

# 2nd wave
# has free pass loaded
df_both %>% filter(wave=="Second") %>%  count(q111, age_groups_control_treat)

```


## New analysis only (2nd wave only)