---
title: "Mobilizing Justice: Montreal Fair Pass Pilot"
author:
  - name: João Pedro Figueira Amorim Parga
    orcid: 0000-0002-4105-5927
    email: joao.parga@mail.utoronto.ca
    affiliations:
      - name: Mobilizing Justice Partnership
format:
  docx: 
    reference-doc: custom-reference-doc.docx
    toc: false
    toc-depth: 2
    number-sections: true
  html:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: true
    code_folding: hide
    theme: readable
    embed-resources: true
  pdf:
    fig-width: 6
    fig-asp: 0.618
    number-sections: true
    colorlinks: true
    keeptex: true
    include-in-header: 
      text: |
        \usepackage{booktabs}
        \usepackage{siunitx}
        \newcolumntype{d}{S[
            input-open-uncertainty=,
            input-close-uncertainty=,
            parse-numbers = false,
            table-align-text-pre=false,
            table-align-text-post=false
         ]}
date: "last-modified"
date-format: '[This version:] MMMM D, YYYY'
abstract: "This version contains only the updated tables and visualizations from the new report. The text can be found at the onedrive link."
csl: csl/american-sociological-association.csl
tbl-cap-location: top
number-sections: true
execute:
  echo: false
  warning: false
  message: false
  cache: false
editor:
  mode: source
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)
# Use cache = TRUE if you want to speed up compilation

knitr::opts_knit$set(output.format = "html")  # Set to "html" for HTML output

# A function to allow for showing some of the inline code
rinline <- function(code){
  html <- '<code  class="r">``` `r CODE` ```</code>'
  sub("CODE", code, html)
}
```

```{r packages}
#| echo: false
#| include: false
library(tidyverse)
library(sf)
library(data.table)
library(janitor)
library(fs)
library(cancensus)
library(tmap)
library(DataExplorer)
library(patchwork)
library(mapview)
library(modelsummary)
library(kableExtra)
library(knitr)
library(gt)
library(gtsummary)
library(conflicted)
library(stars)
library(showtext)
library(sysfonts)
library(hrbrthemes)
library(flextable)
# library(Z.gtsummary.addons)
```

```{r conflicts}
#| include: false
select <- dplyr::select
conflicted::conflicts_prefer(dplyr::filter)
```

```{r fonts}
#| include: false
#| # use different fonts
sysfonts::font_add_google("Roboto", "roboto")
showtext::showtext_auto()
```


```{r}
knitr::opts_chunk$set(tab.topcaption = TRUE)
```


<!-- READ AND MANIPULATE DATA -->

```{r readdata}
#| include: false

load(file = "data/df_both_waves.RData")
load(file = "data/list_count_trips_wave.RData")

```



<!-- START OF THE DOCUMENT -->

# Descriptive analysis


## Who took the survey?

### TAB response rate CONTROL X TREATMENT

```{r}
df_first <- df_both %>% 
  filter(wave=="First")
```

contigency test between response rate and age_control_treatment variable
```{r}
pspearman::spearman.test(
  df_first$response_rate, df_first$age_groups_control_treat
  )
```

```{r tbldescriptiveresponserateage, tab.id="descriptive", tab.cap="Response rate between first and second wave"}
#| echo: false
#| tbl-cap: Response rate between first and second wave

df_both %>% 
  filter(wave=="First") %>% 
  select(
 response_rate,age_groups_control_treat, age_groups_5_all
) %>% 
  rename(
    'Age groups (Control and Treatment)' = age_groups_control_treat
    , "Age groups (5 years intervals)" = age_groups_5_all
  ) %>% 
  mutate(
    `Age groups (5 years intervals)` = fct_recode(
      `Age groups (5 years intervals)`
      , `Age 50-54` = "age_50_54"
      , `Age 55-59` = "age_55_59"
      , `Age 60-64` = "age_60_64"
      , `Age 65-69` = "age_65_69"
      , `Age 70-74` = "age_70_74"
      , `Age 75-79` = "age_75_79"
      , `Age 80-84` = "age_80_84"
      , `Age 85-89` = "age_85_89"
      , `Age 90 and over` = "age_90_over"
      )
  ) %>% 
  st_drop_geometry() %>% 
  gtsummary::tbl_summary(
    type = gtsummary::all_dichotomous() ~ "categorical"
    , by = response_rate
    , sort = list(
      `Age groups (5 years intervals)` ~ "alphanumeric"
      , `Age groups (Control and Treatment)` ~ "frequency"
    ) 
    # all_categorical() ~ "alphanumeric"
    , missing_text = "Missing"
  ) %>% 
  gtsummary::modify_caption(caption = "Age groups of the respondents") %>% 
  as_flex_table() %>% 
  flextable::font(fontname = "Arial")

# gtsummary::as_gt() %>% 
# gt::tab_options(table.font.names = "Arial")
#%>% 
# as_flex_table() 
#%>% 
# gtsummary::modify_header(label = "Demographics") %>% 
# gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables")
```

### TAB response rate DEMOGRAPHICS

contigency test between response rate and age_control_treatment variable
```{r}
pspearman::spearman.test(
  df_first$response_rate, df_first$visible_minority
  )

```

```{r}
pspearman::spearman.test(
  df_first$response_rate, df_first$immigrant
  )
```

```{r tbldescriptiveresponserateinvariable, tab.id="descriptive", tab.cap="Response rate between first and second wave"}
#| echo: false
#| tbl-cap: Response rate between first and second wave

df_both %>% 
  filter(wave == "First") %>% 
  select(
    response_rate
  , visible_minority
  , immigrant
  , gender
  , age_groups_control_treat
  , educ_has_postsec
) %>% 
  rename(
    "Visible minority" = visible_minority
    , "Is Immigrant" = immigrant
    , "Gender" = gender
    , 'Age Groups Control Treat' = age_groups_control_treat
    , "Has post-secondary education" = educ_has_postsec
  ) %>% 
  tbl_strata(
    strata = response_rate
    , .header = "**{strata}, N = {N}**"
    , ~.x %>% 
      gtsummary::tbl_summary(
        type = gtsummary::all_dichotomous() ~ "categorical"
        , sort = list(
          `Visible minority` ~ "frequency"
          , `Is Immigrant` ~ "frequency"
          , `Gender` ~ "frequency"
          , `Has post-secondary education` ~ "frequency"
        )
        # all_categorical() ~ "frequency"
        , missing_text = "Missing"
        , by = `Age Groups Control Treat`
      ) %>% 
      gtsummary::add_overall(last = T, col_label = "**Total**, N = {N}") %>%
      gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables")
  )  %>% 
  as_flex_table() %>% 
  flextable::font(fontname = "Arial")

# gtsummary::as_gt() %>% 
# gt::tab_options(table.font.names = "Arial")
#%>% 
# as_flex_table() 
#%>% 
# gtsummary::modify_header(label = "Demographics") %>% 
# gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables")
```

<!-- ### TAB age -->

<!-- This subsection contains a general description of who took the first wave of the survey... -->

<!-- ```{r tbldescriptiveage, tab.id="descriptive", tab.cap="Descriptive statistics for categorical age groups"} -->
<!-- #| echo: false -->
<!-- #| tbl-cap: Descriptive statistics for categorical variables -->

<!-- df_both %>% select( -->
<!--  age_groups_control_treat, age_groups_5_all -->
<!-- ) %>%  -->
<!--   rename( -->
<!--     'Age groups (Control and Treatment)' = age_groups_control_treat -->
<!--     , "Age groups (5 years intervals)" = age_groups_5_all -->
<!--   ) %>%  -->
<!--   mutate( -->
<!--     `Age groups (5 years intervals)` = fct_recode( -->
<!--       `Age groups (5 years intervals)` -->
<!--       , `Age 50-54` = "age_50_54" -->
<!--       , `Age 55-59` = "age_55_59" -->
<!--       , `Age 60-64` = "age_60_64" -->
<!--       , `Age 65-69` = "age_65_69" -->
<!--       , `Age 70-74` = "age_70_74" -->
<!--       , `Age 75-79` = "age_75_79" -->
<!--       , `Age 80-84` = "age_80_84" -->
<!--       , `Age 85-89` = "age_85_89" -->
<!--       , `Age 90 and over` = "age_90_over" -->
<!--       ) -->
<!--   ) %>%  -->
<!--   st_drop_geometry() %>%  -->
<!--   gtsummary::tbl_summary( -->
<!--     type = gtsummary::all_dichotomous() ~ "categorical" -->
<!--     , sort = list( -->
<!--       `Age groups (5 years intervals)` ~ "alphanumeric" -->
<!--       , `Age groups (Control and Treatment)` ~ "frequency" -->
<!--     )  -->
<!--     # all_categorical() ~ "alphanumeric" -->
<!--     , missing_text = "Missing" -->
<!--   ) %>%  -->
<!--   gtsummary::modify_caption(caption = "Age groups of the respondents") %>%  -->
<!--   as_flex_table() %>%  -->
<!--   flextable::font(fontname = "Arial") -->

<!--   # gtsummary::as_gt() %>%  -->
<!--   # gt::tab_options(table.font.names = "Arial") -->
<!-- #%>%  -->
<!--   # as_flex_table()  -->
<!-- #%>%  -->
<!--   # gtsummary::modify_header(label = "Demographics") %>%  -->
<!--   # gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables") -->
<!-- ``` -->
<!-- ### Table invariable characteristics through time -->

<!-- ```{r tbldescriptiveinvariable, tab.id="descriptive", tab.cap="Descriptive statistics for categorical variables constant over time"} -->
<!-- #| echo: false -->
<!-- #| tbl-cap: Descriptive statistics for categorical variables constant over time -->

<!-- df_both %>% filter(wave=="Second") %>%  -->
<!--   select( -->
<!--   visible_minority -->
<!--   , immigrant -->
<!--   , gender -->
<!--   , age_groups_control_treat -->
<!--   , educ_has_postsec -->
<!-- ) %>%  -->
<!--   rename( -->
<!--     "Visible minority" = visible_minority -->
<!--     , "Is Immigrant" = immigrant -->
<!--     , "Gender" = gender -->
<!--     , 'Age Groups Control Treat' = age_groups_control_treat -->
<!--     , "Has post-secondary education" = educ_has_postsec -->
<!--   ) %>%  -->
<!--   gtsummary::tbl_summary( -->
<!--     type = gtsummary::all_dichotomous() ~ "categorical" -->
<!--     , sort = list( -->
<!--       `Visible minority` ~ "frequency" -->
<!--       , `Is Immigrant` ~ "frequency" -->
<!--       , `Gender` ~ "frequency" -->
<!--       , `Has post-secondary education` ~ "frequency" -->
<!--     ) -->
<!--       # all_categorical() ~ "frequency" -->
<!--     , missing_text = "Missing" -->
<!--     , by = `Age Groups Control Treat` -->
<!--   ) %>%  -->
<!--   gtsummary::add_overall(last = T, col_label = "**Total**, N = {N}") %>% -->
<!--   gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables") %>%  -->
<!--   as_flex_table() %>%  -->
<!--   flextable::font(fontname = "Arial") -->

<!--   # gtsummary::as_gt() %>%  -->
<!--   # gt::tab_options(table.font.names = "Arial") -->
<!-- #%>%  -->
<!--   # as_flex_table()  -->
<!-- #%>%  -->
<!--   # gtsummary::modify_header(label = "Demographics") %>%  -->
<!--   # gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables") -->
<!-- ``` -->




### Table time-variant socio-demographics


```{r tbltimevariantdemographics, tab.id="descriptive", tab.cap="Descriptive statistics for categorical variables variable that vary over time"}
#| echo: false
#| tbl-cap: Descriptive statistics for categorical variables variable that vary over time

df_both %>% filter(response_rate=="Started wave 2") %>% 
  select(
  wave
  , age_groups_control_treat
  , any_mobility_limitation
  , any_functional_limitation
  , employment
  , income_groups
  , in_lim
  , lone_household
  , has_vehicle_household
) %>% 
  rename(
    'Age Groups Control Treat' = age_groups_control_treat
    # , "Age groups" = age_groups_5_all
    , "Any mobility limitation" = any_mobility_limitation
    , "Any functional limitation" = any_functional_limitation
    , "Employment" = employment
    , "Annual income groups" = income_groups
    , "Household below LIM" = in_lim
    , "Single-person household" = lone_household
    , "Has Vehicle in Household" = has_vehicle_household
  ) %>% 
  tbl_strata(
    strata = wave
    , .header = "**{strata}**"
    , ~.x %>% 
      gtsummary::tbl_summary(
        type = gtsummary::all_dichotomous() ~ "categorical"
        , sort = list(
          `Any mobility limitation` ~ "frequency"
          , `Any functional limitation` ~ "frequency"
          , `Employment` ~ "frequency"
          , `Annual income groups` ~ "alphanumeric"
          , `Household below LIM` ~ "frequency"
          , `Single-person household` ~ "frequency"
          , `Has Vehicle in Household` ~ "frequency"
        )
        # all_categorical() ~ "frequency"
        , missing_text = "Missing"
        , by = `Age Groups Control Treat`
      ) %>% 
      gtsummary::add_overall(last = T, col_label = "**Total**, N = {N}") %>%
      gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables")
  )  %>% 
  as_flex_table() %>% 
  flextable::font(fontname = "Arial")

# gtsummary::as_gt() %>% 
# gt::tab_options(table.font.names = "Arial")
#%>% 
# as_flex_table() 
#%>% 
# gtsummary::modify_header(label = "Demographics") %>% 
# gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables")
```

, any_mobility_limitation
  , any_functional_limitation
  , employment
  , income_groups
  , in_lim
  , lone_household
  

## What is their travel behavior?


```{r tbltravelbehaviour, tab.id="descriptive", tab.cap="Descriptive statistics for categorical variables variable that vary over time"}
#| echo: false
#| tbl-cap: Descriptive statistics for categorical variables variable that vary over time

df_both %>% filter(response_rate=="Started wave 2") %>% 
  select(
  wave
  , age_groups_control_treat
  , any_mobility_limitation
  , any_functional_limitation
  , employment
  , income_groups
  , in_lim
  , lone_household
  , has_vehicle_household
) %>% 
  rename(
    'Age Groups Control Treat' = age_groups_control_treat
    # , "Age groups" = age_groups_5_all
    , "Any mobility limitation" = any_mobility_limitation
    , "Any functional limitation" = any_functional_limitation
    , "Employment" = employment
    , "Annual income groups" = income_groups
    , "Household below LIM" = in_lim
    , "Single-person household" = lone_household
    , "Has Vehicle in Household" = has_vehicle_household
  ) %>% 
  tbl_strata(
    strata = wave
    , .header = "**{strata}**"
    , ~.x %>% 
      gtsummary::tbl_summary(
        type = gtsummary::all_dichotomous() ~ "categorical"
        , sort = list(
          `Any mobility limitation` ~ "frequency"
          , `Any functional limitation` ~ "frequency"
          , `Employment` ~ "frequency"
          , `Annual income groups` ~ "alphanumeric"
          , `Household below LIM` ~ "frequency"
          , `Single-person household` ~ "frequency"
          , `Has Vehicle in Household` ~ "frequency"
        )
        # all_categorical() ~ "frequency"
        , missing_text = "Missing"
        , by = `Age Groups Control Treat`
      ) %>% 
      gtsummary::add_overall(last = T, col_label = "**Total**, N = {N}") %>%
      gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables")
  )  %>% 
  as_flex_table() %>% 
  flextable::font(fontname = "Arial")

# gtsummary::as_gt() %>% 
# gt::tab_options(table.font.names = "Arial")
#%>% 
# as_flex_table() 
#%>% 
# gtsummary::modify_header(label = "Demographics") %>% 
# gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables")
```


### Free pass and barriers 

```{r}
df_both <- df_both %>% 
  mutate(
    has_free_pass = as_factor(case_when(
      wave=="First" & 
        q97 == "Yes" |
        wave=="Second" & q111 == "Yes"
        ~ "Yes"
      , wave=="First" &
        q97 %in% c("No, but I’m not planning to get it", "No, but I’m planning to get it soon") | 
        wave=="Second" & q111 == "No"
        ~ "No"
      , .default = NA_character_
    ))
  ) 
```


```{r dfrename}
#| echo: false
df_rename <- df_both 
data.table::setDT(df_rename)
# df_rename[
#   age_groups_control_treat == "Control (50-64)"
#   , `:=`(has_free_pass_wave1 = NA_character_
#       , has_free_pass_wave2 = NA_character_)
# ]

names(df_rename) <- gsub("_", " ", colnames(df_both)) %>% 
  str_to_title()

```

```{r tbltravelbehaviourold, tab.id="descriptive", tab.cap="Descriptive statistics for travel behaviour variables"}
#| echo: false
#| tbl-cap: Descriptive statistics for travel behaviour variables

df_rename %>% filter(`Response Rate`=="Started wave 2") %>% 
  select(
  Wave
  , `Age Groups Control Treat`
  , `Has Vehicle Household`
  , `Has Free Pass`
  # , `Has Free Pass Wave1`
  # , `Has Free Pass Wave2`
  , `Rides Transit As Would Like`
  , `Has Cost Prevented Transit`
  , `Difficulty Transport Expenses`  
) %>% 
  rename(
    , `Has Vehicle Household?` = `Has Vehicle Household`
    , `Has Free Pass?` = `Has Free Pass`
    , `Rides Transit As Would Like?` = `Rides Transit As Would Like`
    , `Has Cost Prevented Transit?` = `Has Cost Prevented Transit`
    , `Difficulty Paying Transport Expenses?` = `Difficulty Transport Expenses`
    # , `Has Free Pass Wave I` = `Has Free Pass Wave1`
    # , `Has Free Pass Wave II` = `Has Free Pass Wave2`
  ) %>% 
  tbl_strata(
    strata = Wave
    , .header = "**{strata}**"
    , ~.x %>% 
      gtsummary::tbl_summary(
        type = gtsummary::all_dichotomous() ~ "categorical"
        , sort = all_categorical() ~ "frequency"
        , missing_text = "Missing"
        , by = `Age Groups Control Treat`
      ) %>% 
      gtsummary::add_overall(last = T, col_label = "**Total**, N = {N}") %>%
      gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables")
  )  %>% 
  as_flex_table() %>% 
  flextable::font(fontname = "Arial")

# gtsummary::as_gt() %>% 
# gt::tab_options(table.font.names = "Arial")
#%>% 
# as_flex_table() 
#%>% 
# gtsummary::modify_header(label = "Demographics") %>% 
# gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables")
```

## Satisfaction

```{r tblsatisfaction, tab.id="descriptive", tab.cap="Descriptive statistics for satisfaction"}
#| echo: false
#| tbl-cap: Descriptive statistics for satisfaction

df_rename %>% filter(`Response Rate`=="Started wave 2") %>%
  select(
    `Wave`
    , `Age Groups Control Treat`
    , `Satisfied Transit`
    # , `Confident Transit`
    , `Satisfaction Community Events`
    , `Satisfaction Recreation`
    , `Satisfaction Seeing Friends Family`
    , `Satisfaction Activity Participation`
    , `Could Stop Driving`
    # , `Has Free Pass`
    # , `Has Free Pass Wave1`
    # , `Has Free Pass Wave2`
    # , `Rides Transit As Would Like`
    # , `Has Cost Prevented Transit`
    # , `Difficulty Transport Expenses`  
    # , `Count Dest`
    # , `Count Modes`
    # , `Count Trips Total`
    # , `Count Trips Wave`
    # , `Count Trips Type Transit`
  ) %>% 
  tbl_strata(
    strata = `Age Groups Control Treat`
    , .header = "**{strata}**"
    , ~.x %>% 
      gtsummary::tbl_summary(
        type = gtsummary::all_dichotomous() ~ "categorical"
        , sort = all_categorical() ~ "frequency"
        , missing_text = "Missing"
        , by = `Wave`
      ) %>% 
      gtsummary::add_overall(last = T, col_label = "**Total**, N = {N}") %>%
      gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables") %>% 
      add_difference(everything() ~ "smd") %>%
      # add_SMD(location = "both", ci = TRUE, decimals = 2) %>%
      add_p(
        # test = all_continuous() ~ "t.test"
        , test.args = all_tests("chisq.test") ~ list(correct = TRUE)
      ) %>%
      # add a header to the statistic column, which is hidden by default
      # adding the header will also unhide the column
      modify_header() %>%
      modify_fmt_fun(statistic ~ style_sigfig)
  )

```

## trip behaviour

```{r tbltripbehaviour, tab.id="descriptive", tab.cap="Descriptive statistics for trip behaviour"}
#| echo: false
#| tbl-cap: Descriptive statistics for trip behaviour

df_rename %>% filter(`Response Rate`=="Started wave 2") %>%
  select(
    `Wave`
    , `Age Groups Control Treat`
    # , `Satisfied Transit`
    # # , `Confident Transit`
    # , `Satisfaction Community Events`
    # , `Satisfaction Recreation`
    # , `Satisfaction Seeing Friends Family`
    # , `Satisfaction Activity Participation`
    # , `Could Stop Driving`
    # , `Has Free Pass`
    # , `Has Free Pass Wave1`
    # , `Has Free Pass Wave2`
    # , `Rides Transit As Would Like`
    # , `Has Cost Prevented Transit`
    # , `Difficulty Transport Expenses`  
    , `Count Trips Wave`
    , `Avg Trips Weekly Total`
    , `Count Dest`
    # , `Count Modes`
    # , `Count Modes`
    # , `Count Trips Total`
    , `Count Trips Type Transit`
    , `Count Trips Type Car`
    , `Count Trips Type Active`
  ) %>% 
  tbl_strata(
    strata = `Age Groups Control Treat`
    , .header = "**{strata}**"
    , ~.x %>% 
      gtsummary::tbl_summary(
        type = gtsummary::all_dichotomous() ~ "categorical"
        , sort = all_categorical() ~ "frequency"
        , missing_text = "Missing"
        , by = `Wave`
      ) %>% 
      gtsummary::add_overall(last = T, col_label = "**Total**, N = {N}") %>%
      gtsummary::modify_caption(caption = "Descriptive statistics for categorical variables") %>% 
      add_difference(everything() ~ "smd") %>%
      # add_SMD(location = "both", ci = TRUE, decimals = 2) %>%
      add_p(
        # test = all_continuous() ~ "t.test"
        , test.args = all_tests("chisq.test") ~ list(correct = TRUE)
      ) %>%
      # add a header to the statistic column, which is hidden by default
      # adding the header will also unhide the column
      modify_header() %>%
      modify_fmt_fun(statistic ~ style_sigfig)
  )

```

CONTINUAR AQUI: INCLUIR SATISFACTION

depois:
- number of transit trips (difference)
- diff in trips overall
- % people agree stop driving

t test?

# Update

## Old analysis (1st wave only (demographics?))

```{r}
# loaded free pass on card (treat and control)
df_both %>% filter(wave=="Second") %>%  count(q111, age_groups_control_treat)
```

df_both %>% filter(wave=="Second") %>%  count(q111, age_groups_control_treat)


## Update old (1st and 2nd wave combined)

Awareness of free pass and how did that translate to its use
```{r}
# 1st wave
df_both %>% filter(wave=="Second") %>%  count(awereness_free_pass, age_groups_control_treat)
df_both %>% filter(wave=="Second") %>%  count(has_free_pass_wave1, age_groups_control_treat)
df_both %>% filter(wave=="Second") %>%  count(has_free_pass_wave1_categories, age_groups_control_treat)

# type of fare (available for 1st wave only)
df_both %>% filter(wave=="Second" & has_free_pass_wave1 == "Yes") %>%  count(fare_type_before_pass_q98, age_groups_control_treat)
df_both %>% filter(wave=="Second" & has_free_pass_wave1 == "Yes") %>%  count(fare_type_before_pass_q99, age_groups_control_treat)

# 2nd wave
# has free pass loaded
df_both %>% filter(wave=="Second") %>%  count(q111, age_groups_control_treat)

```


## New analysis only (2nd wave only)


# model

```{r}
free_1 <- df_both %>% filter(wave=="First"&age_groups_control_treat=="Treatment (65-over)"&has_free_pass_wave1=="Yes")
free_no_1 <- df_both %>% filter(wave=="First"&age_groups_control_treat=="Treatment (65-over)"&has_free_pass_wave1=="No")
free_2 <- df_both %>% filter(wave=="Second"&age_groups_control_treat=="Treatment (65-over)"&has_free_pass_wave2=="Yes")
free_no_2 <- df_both %>% filter(wave=="Second"&age_groups_control_treat=="Treatment (65-over)"&has_free_pass_wave2=="No")

yes_both <- free_1 %>% filter(response_id%in%free_2$response_id)
no_both <- free_no_1 %>% filter(response_id%in%free_no_2$response_id)
```



