---
title: "Mobilizing Justice: Montreal Fair Pass Pilot"
author:
  - name: João Pedro Figueira Amorim Parga
    orcid: 0000-0002-4105-5927
    email: joao.parga@mail.utoronto.ca
    affiliations:
      - name: Mobilizing Justice Partnership
format:
  docx: 
    reference-doc: custom-reference-doc.docx
    toc: false
    toc-depth: 2
    number-sections: true
  html:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: true
    code_folding: hide
    theme: readable
    embed-resources: true
  pdf:
    fig-width: 6
    fig-asp: 0.618
    number-sections: true
    colorlinks: true
    keeptex: true
    include-in-header: 
      text: |
        \usepackage{booktabs}
        \usepackage{siunitx}
        \newcolumntype{d}{S[
            input-open-uncertainty=,
            input-close-uncertainty=,
            parse-numbers = false,
            table-align-text-pre=false,
            table-align-text-post=false
         ]}
date: "last-modified"
date-format: '[This version:] MMMM D, YYYY'
abstract: "This version contains only the updated tables and visualizations from the new report. The text can be found at the onedrive link."
csl: csl/american-sociological-association.csl
tbl-cap-location: top
number-sections: true
execute:
  echo: false
  warning: false
  message: false
  cache: false
editor:
  mode: source
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)
# Use cache = TRUE if you want to speed up compilation

knitr::opts_knit$set(output.format = "html")  # Set to "html" for HTML output

# A function to allow for showing some of the inline code
rinline <- function(code){
  html <- '<code  class="r">``` `r CODE` ```</code>'
  sub("CODE", code, html)
}
```

```{r packages}
#| echo: false
#| include: false
library(tidyverse)
library(sf)
library(data.table)
library(janitor)
library(fs)
library(cancensus)
library(tmap)
library(DataExplorer)
library(patchwork)
library(mapview)
library(modelsummary)
library(kableExtra)
library(knitr)
library(gt)
library(gtsummary)
library(conflicted)
library(stars)
library(showtext)
library(sysfonts)
library(hrbrthemes)
library(flextable)
library(rstatix)
library(openxlsx)
library(MASS)
library(rcompanion)
library(mvoutlier)
library(psych)
# library(Z.gtsummary.addons)

`%nin%` = Negate(`%in%`)

`%nlike%` = Negate(`%like%`)
```

```{r conflicts}
#| include: false
select <- dplyr::select
conflicted::conflicts_prefer(dplyr::filter)
```

```{r fonts}
#| include: false
#| # use different fonts
sysfonts::font_add_google("Roboto", "roboto")
showtext::showtext_auto()
```


```{r}
knitr::opts_chunk$set(tab.topcaption = TRUE)
```


# Read data

```{r}
load(file="data/df_both_outliers.RData")
load(file="data/l_outcomes.RData")
```


# Filter treatment

<!-- EXAMINE INDIVIDUALS WITH ALL NAS ON GROUPING -->
<!-- ```{r} -->
<!-- 666666 -->
<!-- df_second <- df_both %>%  -->
<!--   filter(wave == "Second") %>%  -->
<!--   mutate( -->
<!--     treatment_50_59 = case_when( -->
<!--       age_corrected >= 65L & has_free_pass_wave2 == "Yes" ~ "Treatment" -->

<!--       , (age_corrected >= 50L & -->
<!--         age_corrected < 60L) & -->
<!--         has_free_pass_wave2 == "No" ~ "Control" -->

<!--       , .default = NA_character_ -->
<!--     ) -->
<!--   ) -->

<!-- df_first <- df_both %>%  -->
<!--   filter(wave=="First") %>%  -->
<!--   mutate(treatment_50_59 = case_when( -->

<!--     response_id %in% setDT(df_second)[treatment_50_59=="Treatment",response_id] ~ "Treatment" -->

<!--     , response_id %in% setDT(df_second)[treatment_50_59=="Control",response_id] ~ "Control" -->

<!--     , .default = NA_character_ -->
<!--   )) -->

<!-- df_both <- data.table::rbindlist( -->
<!--   list(df_first,df_second) -->
<!-- ) -->

<!-- rm(df_first, df_second) -->
<!-- ``` -->


Remember: always include `filter(!is.na(treatment)) %>%` pipe before operations

```{r}

```

# Descriptive statistics

## Demographics

### Constant

```{r}
t_age_treat <- df_both %>% 
  filter(treatment=="Treatment" & wave =="First") %>%
  select(treatment, age_groups_5_grouped) %>% 
  # mutate(
  #   age_groups_5_grouped = forcats::fct_drop(
  #     age_groups_5_grouped
  #     , only = c("age_50_54","age_55_59", "age_60_64")
  #     )
  # ) %>% 
    mutate(
    across(
      .cols = -one_of("treatment")
      # don't change fct below for newer one
      , ~ forcats::fct_explicit_na(.x)
      )
  ) %>% 
  tbl_summary(
    by = treatment
  )

t_age_control_1 <- df_both %>% 
  filter(treatment=="Control"  & wave =="First") %>%
  select(treatment, age_groups_5_grouped) %>% 
  
  # mutate(
  #   age_groups_5_grouped = forcats::fct_drop(
  #     age_groups_5_grouped
  #     , only = c("age_50_54","age_55_59", "age_60_64")
  #     )
  # ) %>% 
    mutate(
    across(
      .cols = -one_of("treatment")
      # don't change fct below for newer one
      , ~ forcats::fct_explicit_na(.x)
      )
  ) %>% 
  tbl_summary(
    by = treatment
  )

t_age_control_2 <- df_both %>% 
  filter(treatment_60_64=="Control" & wave =="First") %>% 
  select(treatment_60_64, age_groups_5_grouped) %>% 
  
  # mutate(
  #   age_groups_5_grouped = forcats::fct_drop(
  #     age_groups_5_grouped
  #     , only = c("age_50_54","age_55_59")
  #     )
  # ) %>% 
    mutate(
    across(
      .cols = -one_of("treatment_60_64")
      # don't change fct below for newer one
      , ~ forcats::fct_explicit_na(.x)
      )
  ) %>% 
  tbl_summary(
    by = treatment_60_64
  )

t_age_control_3 <- df_both %>% 
  filter(treatment_50_59=="Control" & wave =="First") %>%
  select(treatment_50_59, age_groups_5_grouped) %>% 
   
    mutate(
    across(
      .cols = -one_of("treatment_50_59")
      # don't change fct below for newer one
      , ~ forcats::fct_explicit_na(.x)
      )
  ) %>% 
  tbl_summary(
    by = treatment_50_59
  )

l_age <- list(
  "t_age_treat" = t_age_treat
  , "t_age_control_1" = t_age_control_1
  , "t_age_control_2" = t_age_control_2
  ,"t_age_control_3" = t_age_control_3
)

rm(t_age_treat,t_age_control_1,t_age_control_2,t_age_control_3)

tbl_merge(
  tbls = l_age
  , tab_spanner = c(
    "**Treatment: 65+ and pass**"
    ,"**Control: 65+ and no pass**", "**Control: 60-64**", "**Control: 50-59**"
    )
)
```


```{r}
t_demog_treatment <- df_both %>% 
  mutate(
    treatment = factor(treatment, levels = c("Treatment", "Control"))
  ) %>% 
  dplyr::filter(
    wave == "First" &
      !is.na(treatment)
      # treatment=="Treatment"
    ) %>% 
  select(
    # wave
    treatment
  , visible_minority
  , immigrant
  , gender
  , educ_has_postsec
) %>% 
    mutate(
    across(
      .cols = -one_of("treatment")
      # don't change fct below for newer one
      , ~ forcats::fct_explicit_na(.x)
      )
  ) %>% 
      gtsummary::tbl_summary(
        type = gtsummary::all_dichotomous() ~ "categorical"
        , sort = list(
          visible_minority ~ "frequency"
          , immigrant ~ "frequency"
          , gender ~ "frequency"
          , educ_has_postsec ~ "frequency"
        )
        # all_categorical() ~ "frequency"
        # , missing = "always"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = "treatment"
        # , statistic = all_categorical() ~ "{p}%"
      ) 

t_demog_control_2 <- df_both %>% 
  dplyr::filter(
    wave == "First" &
      treatment_60_64=="Control"
    ) %>% 
  select(
    # wave
    treatment_60_64
  , visible_minority
  , immigrant
  , gender
  , educ_has_postsec
) %>% 
    mutate(
    across(
      .cols = -one_of("treatment_60_64")
      # don't change fct below for newer one
      , ~ forcats::fct_explicit_na(.x)
      )
  ) %>% 
      gtsummary::tbl_summary(
        type = gtsummary::all_dichotomous() ~ "categorical"
        , sort = list(
          visible_minority ~ "frequency"
          , immigrant ~ "frequency"
          , gender ~ "frequency"
          , educ_has_postsec ~ "frequency"
        )
        # all_categorical() ~ "frequency"
        # , missing = "always"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = "treatment_60_64"
        # , statistic = all_categorical() ~ "{p}%"
      )

t_demog_control_3 <- df_both %>% 
  dplyr::filter(
    wave == "First" &
      treatment_50_59=="Control"
    ) %>% 
  select(
    # wave
    treatment_50_59
  , visible_minority
  , immigrant
  , gender
  , educ_has_postsec
) %>% 
    mutate(
    across(
      .cols = -one_of("treatment_50_59")
      # don't change fct below for newer one
      , ~ forcats::fct_explicit_na(.x)
      )
  ) %>% 
      gtsummary::tbl_summary(
        type = gtsummary::all_dichotomous() ~ "categorical"
        , sort = list(
          visible_minority ~ "frequency"
          , immigrant ~ "frequency"
          , gender ~ "frequency"
          , educ_has_postsec ~ "frequency"
        )
        # all_categorical() ~ "frequency"
        # , missing = "always"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = "treatment_50_59"
        # , statistic = all_categorical() ~ "{p}%"
      ) 

l_demog <- list(
  "t_demog_treatment" = t_demog_treatment
  , "t_demog_control_2" = t_demog_control_2
  , "t_demog_control_3" = t_demog_control_3
)

rm(t_demog_treatment,t_demog_control_2,t_demog_control_3)

tbl_merge(
  tbls = l_demog
  , tab_spanner = c(
    "**Control: 65+ and no pass**","**Control: 60-64**", "**Control: 50-59**"
    )
) %>% 
  gtsummary::modify_caption(caption = "Descriptive statistics: constant demographics")
```


### Time-variant

```{r}
t_variant_treatment <- df_both %>% 
  mutate(
    treatment = factor(treatment, levels = c("Treatment", "Control"))
  ) %>% 
  dplyr::filter(
    wave == "First" &
      !is.na(treatment)
      # treatment=="Treatment"
    ) %>% 
  select(
    # wave
    treatment
    , any_mobility_limitation
    # , any_functional_limitation
    , employment
    , income_groups2
    , in_lim
    , has_vehicle_household
) %>% 
    mutate(
    across(
      .cols = -one_of("treatment")
      # don't change fct below for newer one
      , ~ forcats::fct_explicit_na(.x)
      )
  ) %>% 
      gtsummary::tbl_summary(
        type = gtsummary::all_dichotomous() ~ "categorical"
        , sort = list(
          any_mobility_limitation ~ "frequency"
          , employment ~ "frequency"
          , income_groups2 ~ "alphanumeric"
          , in_lim ~ "frequency"
          , has_vehicle_household ~ "frequency"
        )
        # all_categorical() ~ "frequency"
        # , missing = "always"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = treatment
        # , statistic = all_categorical() ~ "{p}%"
      )

t_variant_control_2 <- df_both %>% 
  dplyr::filter(
    wave == "First" &
      treatment_60_64=="Control"
    ) %>% 
  select(
    treatment_60_64
    , any_mobility_limitation
    , employment
    , income_groups2
    , in_lim
    , has_vehicle_household
) %>% 
    mutate(
    across(
      .cols = -one_of("treatment_60_64")
      # don't change fct below for newer one
      , ~ forcats::fct_explicit_na(.x)
      )
  ) %>% 
      gtsummary::tbl_summary(
        type = gtsummary::all_dichotomous() ~ "categorical"
        , sort = list(
          any_mobility_limitation ~ "frequency"
          , employment ~ "frequency"
          , income_groups2 ~ "alphanumeric"
          , in_lim ~ "frequency"
          , has_vehicle_household ~ "frequency"
        )
        # all_categorical() ~ "frequency"
        # , missing = "always"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = "treatment_60_64"
        # , statistic = all_categorical() ~ "{p}%"
      )

t_variant_control_3 <- df_both %>% 
  dplyr::filter(
    wave == "First" &
      treatment_50_59=="Control"
    ) %>% 
  select(
    treatment_50_59
    , any_mobility_limitation
    , employment
    , income_groups2
    , in_lim
    , has_vehicle_household
) %>% 
    mutate(
    across(
      .cols = -one_of("treatment_50_59")
      # don't change fct below for newer one
      , ~ forcats::fct_explicit_na(.x)
      )
  ) %>% 
      gtsummary::tbl_summary(
        type = gtsummary::all_dichotomous() ~ "categorical"
        , sort = list(
          any_mobility_limitation ~ "frequency"
          , employment ~ "frequency"
          , income_groups2 ~ "alphanumeric"
          , in_lim ~ "frequency"
          , has_vehicle_household ~ "frequency"
        )
        # all_categorical() ~ "frequency"
        # , missing = "always"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = "treatment_50_59"
        # , statistic = all_categorical() ~ "{p}%"
      )



l_variant <- list(
  "t_variant_treatment" = t_variant_treatment
  , "t_variant_control_2" = t_variant_control_2
  , "t_variant_control_3" = t_variant_control_3
)

rm(t_variant_treatment,t_variant_control_2,t_variant_control_3)

tbl_merge(
  tbls = l_variant
  , tab_spanner = c(
    "**Control: 65+ and no pass**","**Control: 60-64**", "**Control: 50-59**"
    )
) %>% 
  gtsummary::modify_caption(caption = "Descriptive statistics: time-variant demographics")

```


# Dependent

## basic

```{r}
t_basic_treatment <- df_both %>% 
    dplyr::filter(
    treatment=="Treatment"
    ) %>%
  mutate(
    treatment = factor(
      treatment
      , levels = c("Treatment", "Control")
      , labels = c("Treatment","Control: 65+ and no pass")
      )
  ) %>% 
  select(
# controls
    treatment
    , wave
# trip generation
    , avg_trips_daily_wave
    # , count_trips_wave
    # , count_modes
    # , count_dest
    # , main_dest
    # , main_mode
    , main_type
# travel behavior
    , rides_transit_as_would_like
# satisfaction
    # , q109_3:q109_5
    , satisfaction_activity_participation
    , satisfaction_seeing_friends_family
    , satisfaction_recreation
    , satisfaction_community_events
# cost
    , difficulty_transport_expenses
    , has_cost_prevented_transit
    # too few observations for treatment (not so for treatment 60-64)
    # , transit_barrier_cost_related
    # not enough obs for treament, but might be different for others
    # , when_cost_transport_barrier
) %>% 
  #   mutate(
  #   across(
  #     .cols = -one_of("treatment_50_59")
  #     # don't change fct below for newer one
  #     , ~ forcats::fct_explicit_na(.x)
  #     )
  # ) %>% 
  # rename(
  #   "50-59" = count_trips_wave
  # ) %>% 
  tbl_strata(
    strata = treatment
    , .tbl_fun = 
      ~ .x %>% 
      tbl_summary(
        type = list(
          # count_trips_wave ~ "continuous"
          # , count_modes ~ "continuous"
          # , count_dest ~ "continuous"
           has_cost_prevented_transit ~ "categorical"
          , rides_transit_as_would_like ~ "categorical"
          # , transit_barrier_cost_related ~ "categorical"
        )
        # type = gtsummary::all_continuous() ~ "continuous2"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = "wave"
        ,  statistic = all_continuous() ~ c(
      "{mean} ({sd})"
      # ,"{median} ({p25}, {p75})"
      )
      , sort = list(
        # main_dest ~ "frequency"
        # , main_mode ~ "frequency"
         main_type ~ "frequency"
        , satisfaction_activity_participation ~ "alphanumeric"
        , satisfaction_seeing_friends_family ~ "alphanumeric"
        , satisfaction_recreation ~ "alphanumeric"
        , satisfaction_community_events ~ "alphanumeric"
      )
      ) #%>% 
      # add_n()
  )

t_basic_treatment_1 <- df_both %>% 
    dplyr::filter(
    treatment=="Control"
    ) %>%
  mutate(
    treatment = factor(
      treatment
      , levels = c("Treatment", "Control")
      , labels = c("Treatment","Control: 65+ and no pass")
      )
  ) %>% 
  select(
# controls
    treatment
    , wave
# trip generation
    , avg_trips_daily_wave
    # , count_trips_wave
    # , count_modes
    # , count_dest
    # , main_dest
    # , main_mode
    , main_type
# travel behavior
    , rides_transit_as_would_like
# satisfaction
    # , q109_3:q109_5
    , satisfaction_activity_participation
    , satisfaction_seeing_friends_family
    , satisfaction_recreation
    , satisfaction_community_events
# cost
    , difficulty_transport_expenses
    , has_cost_prevented_transit
    # too few observations for treatment (not so for treatment 60-64)
    # , transit_barrier_cost_related
    # not enough obs for treament, but might be different for others
    # , when_cost_transport_barrier
) %>% 
  #   mutate(
  #   across(
  #     .cols = -one_of("treatment_50_59")
  #     # don't change fct below for newer one
  #     , ~ forcats::fct_explicit_na(.x)
  #     )
  # ) %>% 
  # rename(
  #   "50-59" = count_trips_wave
  # ) %>% 
  tbl_strata(
    strata = treatment
    , .tbl_fun = 
      ~ .x %>% 
      tbl_summary(
        type = list(
          # count_trips_wave ~ "continuous"
          # , count_modes ~ "continuous"
          # , count_dest ~ "continuous"
           has_cost_prevented_transit ~ "categorical"
          , rides_transit_as_would_like ~ "categorical"
          # , transit_barrier_cost_related ~ "categorical"
        )
        # type = gtsummary::all_continuous() ~ "continuous2"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = "wave"
        ,  statistic = all_continuous() ~ c(
      "{mean} ({sd})"
      # ,"{median} ({p25}, {p75})"
      )
      , sort = list(
        # main_dest ~ "frequency"
        # , main_mode ~ "frequency"
         main_type ~ "frequency"
        , satisfaction_activity_participation ~ "alphanumeric"
        , satisfaction_seeing_friends_family ~ "alphanumeric"
        , satisfaction_recreation ~ "alphanumeric"
        , satisfaction_community_events ~ "alphanumeric"
      )
      ) #%>% 
      # add_n()
  )

t_basic_treatment_2 <- df_both %>% 
    dplyr::filter(
    treatment_60_64=="Control"
    ) %>% 
  mutate(
    treatment_60_64 = factor(
      treatment_60_64
      , levels = c("Treatment", "Control")
      , labels = c("Treatment","Control: 60-64")
      )
  ) %>% 
  select(
# controls
    treatment_60_64
    , wave
# trip generation
    , avg_trips_daily_wave
    # , count_trips_wave
    # , count_modes
    # , count_dest
    # , main_dest
    # , main_mode
    , main_type
# travel behavior
    , rides_transit_as_would_like
# satisfaction
    # , q109_3:q109_5
    , satisfaction_activity_participation
    , satisfaction_seeing_friends_family
    , satisfaction_recreation
    , satisfaction_community_events
# cost
    , difficulty_transport_expenses
    , has_cost_prevented_transit
    # too few observations for treatment (not so for treatment 60-64)
    # , transit_barrier_cost_related
    # not enough obs for treament, but might be different for others
    # , when_cost_transport_barrier
) %>% 
  #   mutate(
  #   across(
  #     .cols = -one_of("treatment_50_59")
  #     # don't change fct below for newer one
  #     , ~ forcats::fct_explicit_na(.x)
  #     )
  # ) %>% 
  # rename(
  #   "50-59" = count_trips_wave
  # ) %>% 
  tbl_strata(
    strata = treatment_60_64
    , .tbl_fun = 
      ~ .x %>% 
      tbl_summary(
        type = list(
          # count_trips_wave ~ "continuous"
          # , count_modes ~ "continuous"
          # , count_dest ~ "continuous"
           has_cost_prevented_transit ~ "categorical"
          , rides_transit_as_would_like ~ "categorical"
          # , transit_barrier_cost_related ~ "categorical"
        )
        # type = gtsummary::all_continuous() ~ "continuous2"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = "wave"
        ,  statistic = all_continuous() ~ c(
      "{mean} ({sd})"
      # ,"{median} ({p25}, {p75})"
      )
      , sort = list(
        # main_dest ~ "frequency"
        # , main_mode ~ "frequency"
        # , main_type ~ "frequency"
         satisfaction_activity_participation ~ "alphanumeric"
        , satisfaction_seeing_friends_family ~ "alphanumeric"
        , satisfaction_recreation ~ "alphanumeric"
        , satisfaction_community_events ~ "alphanumeric"
      )
      ) #%>% 
      # add_n()
  )

t_basic_treatment_3 <- df_both %>% 
    dplyr::filter(
    treatment_50_59=="Control"
    ) %>% 
  mutate(
    treatment_50_59 = factor(
      treatment_50_59
      , levels = c("Treatment", "Control")
      , labels = c("Treatment","Control: 50-59")
      )
  ) %>% 
  select(
# controls
    treatment_50_59
    , wave
# trip generation
    , avg_trips_daily_wave
    # , count_trips_wave
    # , count_modes
    # , count_dest
    # , main_dest
    # , main_mode
    , main_type
# travel behavior
    , rides_transit_as_would_like
# satisfaction
    # , q109_3:q109_5
    , satisfaction_activity_participation
    , satisfaction_seeing_friends_family
    , satisfaction_recreation
    , satisfaction_community_events
# cost
    , difficulty_transport_expenses
    , has_cost_prevented_transit
    # too few observations for treatment (not so for treatment 60-64)
    # , transit_barrier_cost_related
    # not enough obs for treament, but might be different for others
    # , when_cost_transport_barrier
) %>% 
  #   mutate(
  #   across(
  #     .cols = -one_of("treatment_50_59")
  #     # don't change fct below for newer one
  #     , ~ forcats::fct_explicit_na(.x)
  #     )
  # ) %>% 
  # rename(
  #   "50-59" = count_trips_wave
  # ) %>% 
  tbl_strata(
    strata = treatment_50_59
    , .tbl_fun = 
      ~ .x %>% 
      tbl_summary(
        type = list(
          # count_trips_wave ~ "continuous"
          # , count_modes ~ "continuous"
          # , count_dest ~ "continuous"
           has_cost_prevented_transit ~ "categorical"
          , rides_transit_as_would_like ~ "categorical"
          # , transit_barrier_cost_related ~ "categorical"
        )
        # type = gtsummary::all_continuous() ~ "continuous2"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = "wave"
        ,  statistic = all_continuous() ~ c(
      "{mean} ({sd})"
      # ,"{median} ({p25}, {p75})"
      )
      , sort = list(
        # main_dest ~ "frequency"
        # , main_mode ~ "frequency"
         main_type ~ "frequency"
      )
      ) #%>% 
      # add_n()
  )

l_trips_basic <- list(
  "t_basic_treatment" = t_basic_treatment
  , "t_basic_treatment_1" = t_basic_treatment_1
  , "t_basic_treatment_2" = t_basic_treatment_2
  , "t_basic_treatment_3" = t_basic_treatment_3
)

rm(t_basic_treatment, t_basic_treatment_1,t_basic_treatment_2,t_basic_treatment_3)


tbl_merge(
  tbls = l_trips_basic
  , tab_spanner = c(
    "**Treatment: 65+ and pass**"
    , "**Control: 65+ and no pass**","**Control: 60-64**", "**Control: 50-59**"
    )
)

```


## trip generation

Average number of trips per age vs distance to threshold

*50-59 threshold*



```{r}
df_both %>% 
  mutate(
    new_treat = case_when(
      treatment_50_59 == "Control" |
        treatment_60_64 == "Control" ~ "Control"
      , treatment == "Treatment" ~ "Treatment"
      , .default = NA_character_
    )
  ) %>% 
  filter(!is.na(new_treat)) %>%
  group_by(age_corrected, wave, new_treat) %>%
  summarise(avg_trips_age = mean(count_trips_wave, na.rm = T)) %>% 
  mutate(
    age_distance_threshold = case_when(
      is.na(age_corrected) ~ NA_integer_
      , .default = age_corrected - 65L
    )
  )  %>% 
  ggplot(
    aes(
      x = age_distance_threshold
      , y = avg_trips_age
      , colour = new_treat
      , group = new_treat
      )
  ) + 
  geom_point() +
  geom_smooth() +
  geom_vline(
    xintercept = 0L
    , linetype = "dashed"
    , alpha = 0.7
    ) + 
  facet_wrap(~wave)
```

```{r}
df_both %>% 
  mutate(
    new_treat = case_when(
      treatment_50_59 == "Control" |
        treatment_60_64 == "Control" ~ "Control"
      , treatment == "Treatment" ~ "Treatment"
      , .default = NA_character_
    )
  ) %>% 
  filter(!is.na(new_treat)) %>%
  group_by(age_corrected, wave, new_treat) %>%
  summarise(avg_trips_age = mean(count_trips_wave, na.rm = T)) %>% 
  mutate(
    age_distance_threshold = case_when(
      is.na(age_corrected) ~ NA_integer_
      , .default = age_corrected - 65L
    )
  )  %>% 
  ggplot(
    aes(
      x = age_distance_threshold
      , y = avg_trips_age
      , colour = new_treat
      , group = new_treat
      )
  ) + 
  geom_point() +
  geom_smooth() +
  geom_vline(
    xintercept = 0L
    , linetype = "dashed"
    , alpha = 0.7
    ) + 
  facet_wrap(~wave)
```


---


The two tables below show the same information but in different format.
Needs refining.

*Mean trips: treatment repeats in column*


```{r}
t_trips_1 <- df_both %>% 
  dplyr::filter(
    !is.na(treatment)
    ) %>%
  select(
    treatment
    , wave
    , count_trips_wave
) %>% 
  #   mutate(
  #   across(
  #     .cols = -one_of("treatment_50_59")
  #     # don't change fct below for newer one
  #     , ~ forcats::fct_explicit_na(.x)
  #     )
  # ) %>% 
  rename(
    "65+ and no pass" = count_trips_wave
  ) %>% 
  tbl_strata(
    strata = treatment
    , .tbl_fun = 
      ~ .x %>% 
      tbl_summary(
        type = gtsummary::all_continuous() ~ "continuous2"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = "wave"
        ,  statistic = all_continuous() ~ c(
      "{mean} ({sd})",
      "{median} ({p25}, {p75})"
      )
      ) %>% 
      add_n() %>% 
      modify_header(
        all_stat_cols() ~ "**{level}**"
        , label = "**Control**"
        )
  )

t_trips_2 <- df_both %>% 
  dplyr::filter(
    !is.na(treatment_60_64)
    ) %>%
  select(
    treatment_60_64
    , wave
    , count_trips_wave
) %>% 
  #   mutate(
  #   across(
  #     .cols = -one_of("treatment_50_59")
  #     # don't change fct below for newer one
  #     , ~ forcats::fct_explicit_na(.x)
  #     )
  # ) %>% 
    rename(
    "60-64" = count_trips_wave
  ) %>% 
  tbl_strata(
    strata = treatment_60_64
    , .tbl_fun = 
      ~ .x %>% 
      tbl_summary(
        type = gtsummary::all_continuous() ~ "continuous2"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = "wave"
        ,  statistic = all_continuous() ~ c(
      "{mean} ({sd})",
      "{median} ({p25}, {p75})"
      )
      ) %>% 
      add_n()
  )
      
t_trips_3 <- df_both %>% 
  dplyr::filter(
    !is.na(treatment_50_59)
    ) %>%
  select(
    treatment_50_59
    , wave
    , count_trips_wave
) %>% 
  #   mutate(
  #   across(
  #     .cols = -one_of("treatment_50_59")
  #     # don't change fct below for newer one
  #     , ~ forcats::fct_explicit_na(.x)
  #     )
  # ) %>% 
  rename(
    "50-59" = count_trips_wave
  ) %>% 
  tbl_strata(
    strata = treatment_50_59
    , .tbl_fun = 
      ~ .x %>% 
      tbl_summary(
        type = gtsummary::all_continuous() ~ "continuous2"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = "wave"
        ,  statistic = all_continuous() ~ c(
      "{mean} ({sd})",
      "{median} ({p25}, {p75})"
      )
      ) %>% 
      add_n()
  )

l_trips <- list(
  "t_trips_1" = t_trips_1
  , "t_trips_2" = t_trips_2
  , "t_trips_3" = t_trips_3
)

rm(t_trips_1,t_trips_2,t_trips_3)

tbl_stack(l_trips)

```

```{r new}
# t_trips_1 <- df_both %>% 
#   dplyr::filter(
#     !is.na(treatment)
#     ) %>%
#     mutate(
#     treatment = factor(treatment, levels = c("Treatment", "Control"))
#   ) %>% 
#   select(
#     treatment
#     , wave
#     , count_trips_wave
# ) %>% 
#   #   mutate(
#   #   across(
#   #     .cols = -one_of("treatment_50_59")
#   #     # don't change fct below for newer one
#   #     , ~ forcats::fct_explicit_na(.x)
#   #     )
#   # ) %>% 
#   rename(
#     "65+ and pass" = count_trips_wave
#   ) %>% 
#   tbl_strata(
#     strata = treatment
#     , .tbl_fun = 
#       ~ .x %>% 
#       tbl_summary(
#         type = gtsummary::all_continuous() ~ "continuous2"
#         , missing_text = "Missing"
#         , missing_stat = "{p_miss}%"
#         , by = "wave"
#         ,  statistic = all_continuous() ~ c(
#       "{mean} ({sd})",
#       "{median} ({p25}, {p75})"
#       )
#       ) %>% 
#       add_n() #%>% 
#       # modify_header(
#       #   all_stat_cols() ~ "**{level}**"
#       #   , label = "**Control**"
#       #   )
#   )

# treatment
t_trips <- df_both %>% 
  dplyr::filter(
    treatment == "Treatment"
    ) %>%
  select(
    # treatment_60_64
    , wave
    , count_trips_wave
) %>% 
  #   mutate(
  #   across(
  #     .cols = -one_of("treatment_50_59")
  #     # don't change fct below for newer one
  #     , ~ forcats::fct_explicit_na(.x)
  #     )
  # ) %>% 
    rename(
    "65+ and pass" = count_trips_wave
  ) %>% 
  tbl_summary(
        type = gtsummary::all_continuous() ~ "continuous2"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = "wave"
        , statistic = all_continuous() ~ c(
      "{mean} ({sd})",
      "{median} ({p25}, {p75})"
      )
      ) 

t_trips_1 <- df_both %>% 
  dplyr::filter(
    treatment == "Control"
    ) %>%
  select(
    # treatment_60_64
    , wave
    , count_trips_wave
) %>% 
  #   mutate(
  #   across(
  #     .cols = -one_of("treatment_50_59")
  #     # don't change fct below for newer one
  #     , ~ forcats::fct_explicit_na(.x)
  #     )
  # ) %>% 
    rename(
    "65+ and no pass" = count_trips_wave
  ) %>% 
  tbl_summary(
        type = gtsummary::all_continuous() ~ "continuous2"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = "wave"
        , statistic = all_continuous() ~ c(
      "{mean} ({sd})",
      "{median} ({p25}, {p75})"
      )
      )

t_trips_2 <- df_both %>% 
  dplyr::filter(
    treatment_60_64 == "Control"
    ) %>%
  select(
    # treatment_60_64
    , wave
    , count_trips_wave
) %>% 
  #   mutate(
  #   across(
  #     .cols = -one_of("treatment_50_59")
  #     # don't change fct below for newer one
  #     , ~ forcats::fct_explicit_na(.x)
  #     )
  # ) %>% 
    rename(
    "60-64" = count_trips_wave
  ) %>% 
  tbl_summary(
        type = gtsummary::all_continuous() ~ "continuous2"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = "wave"
        , statistic = all_continuous() ~ c(
      "{mean} ({sd})",
      "{median} ({p25}, {p75})"
      )
      ) 

t_trips_3 <- df_both %>% 
  dplyr::filter(
    treatment_50_59 == "Control"
    ) %>%
  select(
    # treatment_60_64
    , wave
    , count_trips_wave
) %>% 
  #   mutate(
  #   across(
  #     .cols = -one_of("treatment_50_59")
  #     # don't change fct below for newer one
  #     , ~ forcats::fct_explicit_na(.x)
  #     )
  # ) %>% 
    rename(
    "50-59" = count_trips_wave
  ) %>% 
  tbl_summary(
        type = gtsummary::all_continuous() ~ "continuous2"
        , missing_text = "Missing"
        , missing_stat = "{p_miss}%"
        , by = "wave"
        , statistic = all_continuous() ~ c(
      "{mean} ({sd})",
      "{median} ({p25}, {p75})"
      )
      ) 


l_trips_v2 <- list(
  "t_trips" = t_trips
  ,"t_trips_1" = t_trips_1
  , "t_trips_2" = t_trips_2
  , "t_trips_3" = t_trips_3
)

tbl_merge(
  tbls = l_trips_v2
  , tab_spanner = c(
    "**Treatment: 65+ and pass**"
    , "**Control: 65+ and no pass**","**Control: 60-64**", "**Control: 50-59**"
    )
)

```


riders enough transit:
```{r}
# fisher test (2x2)
stats::fisher.test(
  df_outcome_cost$treatment
  , df_outcome_cost$delta_rides_enough_transit
)

stats::fisher.test(
  df_outcome_cost$treatment
  , df_outcome_cost$delta_rides_enough_transit
)

rstatix::cramer_v(
  df_outcome_cost$treatment
  , df_outcome_cost$delta_rides_enough_transit
)
# post-hoc test: assess if the differences between levels was significant
rstatix::pairwise_fisher_test(
  xtab = table(
    df_outcome_cost$treatment
    ,df_outcome_cost$delta_rides_enough_transit
    )
  , p.adjust.method = "bonferroni"
)

```


## cost

```{r}

```


## activity participation

### conbrach-alpha

Check if activity participation variables can be grouped together

Check if the variables can actually be grouped together -> cronbach alpha
https://rforhr.com/cronbachsalpha.html

Raw alpha of 0.85 suggest that it is above the threshold of internal consistency.
Moreover, the alpha does not increase when dropping any of the variables (Reliability if an item is dropped), thus they seem to be consistent with one another.

```{r}
first <- df_both %>% 
  filter(wave=="First") %>% 
  select(q109_3:q109_5)

# turn variables into numerical (so that the alpha funciton can run)
psych::alpha(psych::char2numeric(first))

```

### Mann-Whitney test

https://guides.library.lincoln.ac.uk/mash/mann_whitney_R

```{r}

```

